Zaca Fire Type Conversion

### 2020-08-17 Model Data Sheets and Code/Zaca MCF Mortality Model Results.Rmd
vsurf and model selection

mcf <- read.csv('Zaca MCF variables for model.csv') %>% 
  mutate(mortality_qual= fct_relevel(mortality_qual, levels=c("High", "Moderate", "Low", "None"))) %>% 
  mutate(dnbr_qualitative = case_when(
    dnbr_2007 < 0.1 ~ "None",
    dnbr_2007 < 0.27 ~ "Low",
    dnbr_2007 < 0.66 ~ "Moderate",
    dnbr_2007 > 0.66 ~ "High")) %>%
  mutate(dnbr_qualitative = as.factor(dnbr_qualitative),
         dnbr_qualitative = fct_relevel(dnbr_qualitative, levels=c("High", "Moderate", "Low", "None"))) %>%
  mutate(aspect_transformed = (1-cos((2*pi*aspect)/360))/2) %>% 
  mutate(three_class_mortality = case_when(
    mortality_qual %in% "High" ~ "High",
    mortality_qual %in% "Moderate" ~ "Moderate",
    mortality_qual %in% c("Low", "None") ~ "Low")) %>% 
  mutate(three_class_mortality = as.factor(three_class_mortality)) %>% 
  mutate(three_class_mortality= fct_relevel(three_class_mortality, 
                                     levels=c("Low", "Moderate", "High"))) %>% 
  mutate(two_class_mortality = case_when(
    mortality_qual %in% "High" ~ "High",
    mortality_qual %in% c("Moderate", "Low", "None") ~ "Other")) %>% 
  mutate(two_class_mortality = as.factor(two_class_mortality)) %>% 
  mutate(two_class_mortality= fct_relevel(two_class_mortality, 
                                     levels=c("Other", "High"))) %>% 
  mutate(tc_mortality = as.factor(tc_mortality)) %>% 
  mutate(rdnbr_qualitative = case_when(
    rdnbr_2007 < 69 ~ "None",
    rdnbr_2007 < 315 ~ "Low",
    rdnbr_2007 < 640 ~ "Moderate",
    rdnbr_2007 > 640 ~ "High")) %>%
  mutate(rdnbr_qualitative = as.factor(rdnbr_qualitative),
         rdnbr_qualitative = fct_relevel(rdnbr_qualitative, 
                                         levels=c("High", "Moderate", "Low", "None"))) %>% 
    mutate(rdnbr_qualitative2 = case_when(
    rdnbr_qualitative %in% "High" ~ "High",
    rdnbr_qualitative %in% c("Moderate", "Low", "None") ~ "Other")) %>% 
  mutate(rdnbr_qualitative2 = as.factor(rdnbr_qualitative2)) %>% 
  mutate(rdnbr_qualitative2= fct_relevel(rdnbr_qualitative2, 
                                     levels=c("Other", "High")))

@@

# mcf type conversison ------------------------------------------------------------------- 
tc_count <- mcf %>% 
  group_by(tc_mortality) %>% 
  count() # 99 stands with type conversion, 636 with no type conversion

@@

>>>>

mcf <- read.csv('Zaca MCF variables for model.csv') %>% 
  mutate(mortality_qual= fct_relevel(mortality_qual, levels=c("High", "Moderate", "Low", "None"))) %>% 
  mutate(dnbr_qualitative = case_when(
    dnbr_2007 < 0.1 ~ "None",
    dnbr_2007 < 0.27 ~ "Low",
    dnbr_2007 < 0.66 ~ "Moderate",
    dnbr_2007 > 0.66 ~ "High")) %>%
  mutate(dnbr_qualitative = as.factor(dnbr_qualitative),
         dnbr_qualitative = fct_relevel(dnbr_qualitative, levels=c("High", "Moderate", "Low", "None"))) %>%
  mutate(aspect_transformed = (1-cos((2*pi*aspect)/360))/2) %>% 
  mutate(three_class_mortality = case_when(
    mortality_qual %in% "High" ~ "High",
    mortality_qual %in% "Moderate" ~ "Moderate",
    mortality_qual %in% c("Low", "None") ~ "Low")) %>% 
  mutate(three_class_mortality = as.factor(three_class_mortality)) %>% 
  mutate(three_class_mortality= fct_relevel(three_class_mortality, 
                                     levels=c("Low", "Moderate", "High"))) %>% 
  mutate(two_class_mortality = case_when(
    mortality_qual %in% "High" ~ "High",
    mortality_qual %in% c("Moderate", "Low", "None") ~ "Other")) %>% 
  mutate(two_class_mortality = as.factor(two_class_mortality)) %>% 
  mutate(two_class_mortality= fct_relevel(two_class_mortality, 
                                     levels=c("Other", "High"))) %>% 
  mutate(tc_mortality = as.factor(tc_mortality)) %>% 
  mutate(rdnbr_qualitative = case_when(
    rdnbr_2007 < 69 ~ "None",
    rdnbr_2007 < 315 ~ "Low",
    rdnbr_2007 < 640 ~ "Moderate",
    rdnbr_2007 > 640 ~ "High")) %>%
  mutate(rdnbr_qualitative = as.factor(rdnbr_qualitative),
         rdnbr_qualitative = fct_relevel(rdnbr_qualitative, 
                                         levels=c("High", "Moderate", "Low", "None"))) %>% 
    mutate(rdnbr_qualitative2 = case_when(
    rdnbr_qualitative %in% "High" ~ "High",
    rdnbr_qualitative %in% c("Moderate", "Low", "None") ~ "Other")) %>% 
  mutate(rdnbr_qualitative2 = as.factor(rdnbr_qualitative2)) %>% 
  mutate(rdnbr_qualitative2= fct_relevel(rdnbr_qualitative2, 
                                     levels=c("Other", "High")))  %>% 
  mutate(rdnbr_tc = case_when(
    rdnbr_2007 < 1000 ~ "Not_TC",
    rdnbr_2007 >= 1000 ~ "TC"
  )) %>% 
  mutate(rdnbr_tc = as.factor(rdnbr_tc)) %>% 
  mutate(rdnbr_tc= fct_relevel(rdnbr_tc, 
                                     levels=c("Not_TC", "TC")))

@@

# mcf manual type conversison ------------------------------------------------------------------- 
tc_count <- mcf %>% 
  group_by(tc_mortality) %>% 
  count() # 99 stands with type conversion, 636 with no type conversion

# mcf type conversion with rdnbr > 1000 threshold --------------------------------------------
tc2_coun <- mcf %>% 
  group_by(rdnbr_tc) %>% 
  count() #not tc = 680, tc = 55

@@

### Stepwise Variable Selection

mcf_sub <- mcf %>% 
  dplyr::select(rdnbr_qualitative,flow_accumulation, sr, rain, slope, tmax, TPI_five, TWI, fri, aspect_transformed, cwd, ndvi, vpd_aug, vpd_jan,ndmi_2006, dem_10m_zaca)

#Extract prediction and response variables
mcf.predictors <- mcf_sub[,2:ncol(mcf_sub)]
mcf.response <- mcf_sub[,1]

#Run stepwise variable selection algorithm
mcf.vsurf <- VSURF(x = mcf.predictors, y = mcf.response)

#Get names of selected variables
names(mcf.predictors[mcf.vsurf$varselect.pred])

# does vsurf selected variables improve variance explained and result in stabilized variable importance ranking?

for (i in 1:7) {

rfm <- randomForest(rdnbr_tc ~ rain + tmax + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + ndvi + ndmi_2006 + aspect_transformed + vpd_jan, 
                    data = mcf, 
                    sampsize = c('Not_TC' = 55, 'TC' = 55), 
                    ntree=1500, 
                    importance=T,
                    replace = T)
# OOB and importance
rfm
importance(rfm)
#plot to visualize variable importance
varImpPlot(rfm)

}
rfm_roc <-  roc(mcf$rdnbr_tc, rfm$votes[,1])
auc(rfm_roc)

```
all variables: rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + ndvi + ndmi_2006 + vpd_jan + aspect_transformed

vsurf variable selection (mcf, regression): slope + sr + vpd_aug + tmax + ndmi_2006 + vpd_jan + ndvi + rain + dem_10m_zaca

vsurf variables: % variance explained = 50%, VI ranking is NOT stable, msr = 44,577
all variables: % variance explained = 47%, VI ranking is NOT stable, msr = 47,592
all variables (minus elevation): % variance explained = 47%, VI ranking is mostly stable
all variables (minus vpd_jan and elevation): % variance explained = 44%, VI ranking IS very stable
all variables (minus aspect, tpi, twi): % variance explained = 50%, VI ranking is not that stable, msr = 44,699
all variables (minus aspect, tpi, twi, vpd_jan): % variance explained = 49%, VI ranking is kinda stable
all variables (minus aspect, tpi, twi, elevation): % variance explained = 49%, VI ranking is kinda stable

vsurf variable selection (mcf, classification, rdnr_tc): vpd_aug + slope + rain + vpd_jan + ndvi

vsurf variables: oob = 14%, not tc class error = 13%, tc class error = 25%, VI ranking IS stable, auc = 89%
all variables: oob = 15%, not tc class error = 14%, tc class error = 31%, VI ranking is NOT stable, auc = 86%
all variables (minus elevation, tpi, and vpd_jan): oob = 15%, not tc class error = 14%, tc class error = 35%, VI ranking IS stable, auc = 84%
all variables(minus elevation, tpi): oob = 15%, not tc class error = 13%, tc class error = 31%, VI ranking is not stable, auc = 86%

vsurf variable selection (mcf, classification, rdnbr 4 mortality classes): sr + ndmi_2006 + vdp_aug + dem_10m_zaca + ndvi + vpd_jan + tmax + cwd + rain + flow_accumulation

### Stepwise Variable Selection - PICO

pico_sub <- pico %>% 
  dplyr::select(flow_accumulation, insolation, rain, slope, tmax, TPI_five, TWI,  aspect_transformed, cwd,  vpd_aug, vpd_jan,ndmi_2006, dem, years_since_burn) # no ndvi

#Extract prediction and response variables
pico.predictors <- pico_sub[,c(2:14)]
pico.response <- pico_sub[, 1]

#Run stepwise variable selection algorithm
pico.vsurf <- VSURF(pico$rdnbr_2007 ~ ., data = pico_sub) #setting up the code the same way as the mcf resulted in an error, but this format also works

#Get names of selected variables
names(pico_sub[pico.vsurf$varselect.pred])

# does vsurf selected variables improve variance explained and result in stabilized variable importance ranking?

for (i in 1:7) {

rfm <- randomForest(rdnbr_2007 ~ flow_accumulation + insolation + + rain + ndmi_2006 + tmax + dem + vpd_jan + vpd_aug + cwd + years_since_burn, 
                    data = pico, 
                    #sampsize = c('Not_TC' = 55, 'TC' = 55), 
                    ntree=1500, 
                    importance=T,
                    replace = T)
# OOB and importance
rfm
importance(rfm)
#plot to visualize variable importance
varImpPlot(rfm)

}
rfm_roc <-  roc(mcf$rdnbr_tc, rfm$votes[,1])
auc(rfm_roc)

```
VSURF regreession variables selected: flow accumulation + insolation + + rain + ndmi_2006 + tmax + dem + vpd_jan + vpd_aug + cwd + years_since_burn
